---
title: "Chapter 1 - Exercises"
author: "Héctor Lira Talancón"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package Installation

```{r libraries, message = F}
library(tidyr)
library(dplyr)
library(lubridate)
library(data.table)
library(ggplot2)
library(survival)
```

## Reading the Data

```{r read_data}
whas100 <- fread("Second-Edition Data/whas100.dat",
                 col.names=c("id", "admitdate", "foldate", "los", "lenfol", "fstat", "age", "gender", "bmi")) %>%
  mutate(admitdate=as.Date(admitdate, format='%m/%d/%Y'),
         foldate=as.Date(foldate, format='%m/%d/%Y'),
         fstat=as.factor(fstat)) %>%
  as_tibble
```

## Exercise 1

```{r scatter_plot_1}
ggplot(data=whas100) +
  geom_point(aes(y=foldate, x=age, shape=fstat)) +
  ggtitle("Follow-up time vs. age")
```

### 1.a
Differences from a normal scatter plot in a linear regression setting:

- there are points in the scatter plot that "stop" at the last follow-up date (right censored).
- there are three dimensions in this scatter plot relevant to our regression problem (follow-up date, age, and vital status).

### 1.b
We'll draw a line using local averages.

```{r average_foldate}
cuts <- pretty(whas100$age)
class_reps <- frollmean(cuts, n=2)[-1]

foldate_by_age_group <- whas100 %>%
  mutate(age_group=cut(age,
                       breaks=cuts,
                       include.lowest=T,
                       right=F)) %>%
  group_by(age_group) %>%
  summarise(mean_foldate=mean(foldate)) %>%
  mutate(class_reps=class_reps)

ggplot() +
  geom_point(data=whas100, aes(y=foldate, x=age, shape=fstat)) +
  geom_line(data=foldate_by_age_group, aes(y=mean_foldate, x=class_reps), color="blue") +
  ggtitle("Follow-up time vs. age with regression line")
```

### 1.c
It is not a straight line. We could use a step function to represent this regression line.

### 1.d
It is. R has a packages called `survival` to fit models like these.

## Exercise 2
Follow-up dates may be right censored.

### Exercise 3
Event: drop-out of the study.

Right censored observations include:
- patients that could not be contacted anymore (e.g. they moved cities).

Left censored observations include:
- 




